---
title: "MSD Homework 2, Problem 3"
author: "Mane Diouf & Mary McTarsney"
date: '`r Sys.time()`'
output:
  html_document:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
library(here)
library(scales)
library(tidyverse)
library(cfbfastR)

theme_set(theme_bw())

knitr::opts_chunk$set(echo = TRUE)
```

# Define the football game schedules

```{r}
sched <- map_dfr(2000:2005, espn_cfb_schedule)
```

# Save the schedule file

```{r}
# Optional, save for use in other files
write_csv(sched, file = "game_schedule.csv")
```

# -- Team to agency mapping --

## Load the batch header file
```{r}
# Select for columns relevant for matching ori to college towns
header <- read_csv("nibrs_batch_header_1991_2024.csv") %>%
    filter(year == 2000) %>%
    select(c("ori","city_name","state_abbreviation","agency_indicator")) %>%
    unique()
```

```{r}
# Create a dataset that contains the needed 26 college towns with their
# corresponding state abbreviations, school names, and short school labels

paper_cities <- tribble(
  ~city_name, ~state_abbreviation, ~school, ~school_short,
  "akron", "oh", "the university of akron", "Akron",
  "ames", "ia", "iowa state university", "Iowa State",
  "ann arbor", "mi", "the university of michigan", "Michigan",
  "athens", "oh", "ohio university", "Ohio",
  "austin", "tx", "the university of texas at austin", "Texas",
  "blacksburg", "va", "virginia polytechnic institute and state university", "Virginia Tech",
  "boise", "id", "boise state university", "Boise State",
  "clemson", "sc",  "clemson university", "Clemson",
  "colorado springs", "co",  "united states air force academy", "Air Force",
  "columbia", "sc", "the university of south carolina", "South Carolina",
  "columbus", "oh",  "the ohio state university", "Ohio State",
  "denton", "tx", "the university of north texas", "North Texas",
  "east lansing", "mi", "michigan state university", "Michigan State",
  "fayetteville", "ar", "university of arkansas", "Arkansas",
  "huntington", "wv", "marshall university", "Marshall",
  "iowa city", "ia", "the university of iowa", "Iowa",
  "jonesboro", "ar", "arkansas state university", "Arkansas State",
  "kalamazoo", "mi", "western michigan university", "Western Michigan",
  "lawrence", "ks", "the university of kansas", "Kansas",
  "logan", "ut",  "utah state university", "Utah State",
  "lubbock", "tx", "texas tech university", "Texas Tech",
  "morgantown", "wv", "west virginia university", "West Virginia",
  "moscow", "id", "the university of idaho", "Idaho",
  "mount pleasant", "mi", "eastern michigan university", "Eastern Michigan",
  "murfreesboro", "tn", "middle tennessee state university", "Middle Tennessee",
  "provo", "ut", "brigham young university", "BYU"
)
```

## Overall records with ori, city, state, and agency agency indicator
```{r}
# Match each college town with respective ori
ori_mapping <- left_join(
    paper_cities, 
    header, 
    by=c("city_name","state_abbreviation")
    ) %>% 
    filter(agency_indicator == "city")

# Save for future use in other files
write_csv(ori_mapping, "ori_mapping.csv")
``` 

# Reproduction of table 2 (Distribution of games by day of week)
```{r}
# Initiate the team names with their matching universities
# so that we can easily match if the teams are either
# home teams or away teams  

university_team_lookup <- tribble(
  ~university, ~team_name,
  'the university of akron', 'Zips',
  'iowa state university', 'Cyclones',
  'the university of michigan', 'Wolverines',
  'ohio university', 'Bobcats',
  'the university of texas at austin', 'Longhorns',
  'virginia polytechnic institute and state university', 'Hokies',
  'boise state university', 'Broncos',
  'clemson university', 'Tigers',
  'united states air force academy', 'Falcons',
  'the university of south carolina', 'Gamecocks',
  'the ohio state university', 'Buckeyes',
  'the university of north texas', 'Mean Green',
  'michigan state university', 'Spartans',
  'university of arkansas', 'Razorbacks',
  'marshall university', 'Thundering Herd',
  'the university of iowa', 'Hawkeyes',
  'arkansas state university', 'Red Wolves',
  'western michigan university', 'Broncos2',
  'the university of kansas', 'Jayhawks',
  'utah state university', 'Aggies',
  'texas tech university', 'Red Raiders',
  'west virginia university', 'Mountaineers',
  'the university of idaho', 'Vandals',
  'eastern michigan university', 'Eagles',
  'middle tennessee state university', 'Blue Raiders',
  'brigham young university', 'Cougars'
)
```
# Recreate table 2 that creates a weekday summary of games and offenses 
# Based on the initial schedule file,
# we filter out the home or away team that played daily
```{r}
game_week <- sched %>%
    mutate(day_of_the_week=wday(game_date, label = TRUE, abbr = FALSE))%>% 
    filter(home_team_name %in% university_team_lookup$team_name | away_team_name %in% university_team_lookup$team_name) %>% 
    group_by(day_of_the_week) %>% 
    summarize(
        num_games=n()
    )

# Add an additional row for "Total" to the existing table 
game_week <- bind_rows(
  game_week,
  tibble(day_of_the_week = "Total", num_games = sum(game_week$num_games))
)

game_week
```