---
title: "MSD Homework 2, Problem 3"
author: "Mane Diouf & Mary McTarsney"
date: '`r Sys.time()`'
output:
  html_document:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
library(here)
library(scales)
library(tidyverse)
library(cfbfastR)

theme_set(theme_bw())

knitr::opts_chunk$set(echo = TRUE)
```

# Define the football game schedules

```{r}
sched <- map_dfr(2000:2005, espn_cfb_schedule)
```

# Save the schedule file

```{r}
# Optional, save for use in other files
write_csv(sched, file = "game_schedule.csv")
```

# -- Team to agency mapping --

## Load the batch header file
```{r}
# Select relevant columns for matching ori to college towns
header <- read_csv("nibrs_batch_header_1991_2024.csv") %>%
    filter(year == 2000) %>%
    select(c("ori","city_name","state_abbreviation","agency_indicator")) %>%
    unique()
```

```{r}
# Create a dataset that contains the needed 26 college towns with their
# corresponding state abbreviations, school names, and short school labels

paper_cities <- tribble(
  ~city_name, ~state_abbreviation, ~school, ~school_short,
  "akron", "oh", "the university of akron", "Akron",
  "ames", "ia", "iowa state university", "Iowa State",
  "ann arbor", "mi", "the university of michigan", "Michigan",
  "athens", "oh", "ohio university", "Ohio",
  "austin", "tx", "the university of texas at austin", "Texas",
  "blacksburg", "va", "virginia polytechnic institute and state university", "Virginia Tech",
  "boise", "id", "boise state university", "Boise State",
  "clemson", "sc",  "clemson university", "Clemson",
  "colorado springs", "co",  "united states air force academy", "Air Force",
  "columbia", "sc", "the university of south carolina", "South Carolina",
  "columbus", "oh",  "the ohio state university", "Ohio State",
  "denton", "tx", "the university of north texas", "North Texas",
  "east lansing", "mi", "michigan state university", "Michigan State",
  "fayetteville", "ar", "university of arkansas", "Arkansas",
  "huntington", "wv", "marshall university", "Marshall",
  "iowa city", "ia", "the university of iowa", "Iowa",
  "jonesboro", "ar", "arkansas state university", "Arkansas State",
  "kalamazoo", "mi", "western michigan university", "Western Michigan",
  "lawrence", "ks", "the university of kansas", "Kansas",
  "logan", "ut",  "utah state university", "Utah State",
  "lubbock", "tx", "texas tech university", "Texas Tech",
  "morgantown", "wv", "west virginia university", "West Virginia",
  "moscow", "id", "the university of idaho", "Idaho",
  "mount pleasant", "mi", "eastern michigan university", "Eastern Michigan",
  "murfreesboro", "tn", "middle tennessee state university", "Middle Tennessee",
  "provo", "ut", "brigham young university", "BYU"
)
```

## Overall records with ori, city, state, and agency agency indicator
```{r}
# Match each college town with respective ori
ori_mapping <- left_join(
    paper_cities, 
    header, 
    by=c("city_name","state_abbreviation")
    ) %>% 
    filter(agency_indicator == "city")

# Save for future use in other files
write_csv(ori_mapping, "ori_mapping.csv")
``` 

# REPRODUCING RESULTS FROM THE PAPER

```{r}
# Initiate the team names with their matching universities
# so that we can easily match if the teams are either
# home teams or away teams  

university_team_lookup <- tribble(
  ~university, ~team_name,
  'the university of akron', 'Zips',
  'iowa state university', 'Cyclones',
  'the university of michigan', 'Wolverines',
  'ohio university', 'Bobcats',
  'the university of texas at austin', 'Longhorns',
  'virginia polytechnic institute and state university', 'Hokies',
  'boise state university', 'Broncos',
  'clemson university', 'Tigers',
  'united states air force academy', 'Falcons',
  'the university of south carolina', 'Gamecocks',
  'the ohio state university', 'Buckeyes',
  'the university of north texas', 'Mean Green',
  'michigan state university', 'Spartans',
  'university of arkansas', 'Razorbacks',
  'marshall university', 'Thundering Herd',
  'the university of iowa', 'Hawkeyes',
  'arkansas state university', 'Red Wolves',
  'western michigan university', 'Broncos2',
  'the university of kansas', 'Jayhawks',
  'utah state university', 'Aggies',
  'texas tech university', 'Red Raiders',
  'west virginia university', 'Mountaineers',
  'the university of idaho', 'Vandals',
  'eastern michigan university', 'Eagles',
  'middle tennessee state university', 'Blue Raiders',
  'brigham young university', 'Cougars'
)
```

# Reproducing Appendix Table 1
```{r}
filtered_sched <- sched %>%    
    filter(
        home_team_name %in% university_team_lookup$team_name | 
        away_team_name %in% university_team_lookup$team_name 
    )
```

# Essentially want each row to be a unique day/town combo, including the count of assault, count of vandalism, and
# **TO DO: denote game type (home/away/no game)**

```{r}
# Load the offense data file
offense <- read_csv("offense.csv")

# Create counts for vandalism and assault for each unique ori/date pair
offense_count <- offense %>%
    mutate( offense_type=case_when(
        str_detect(ucr_offense_code,"vandalism") ~ "vandalism",
        TRUE ~ "assault"
    )) %>%
    select(-ucr_offense_code) %>%
    group_by(ori,incident_date,offense_type) %>%
    summarise(offense_count=n()) %>%
    pivot_wider(
        names_from=offense_type,
        values_from=offense_count
    ) %>% 
    replace_na(list(assault=0,vandalism=0)) %>%
    rename(date=incident_date)
```

# Create a table of base dates we are interested in: Aug 20 - Dec 10 from 2000 to 2005
```{r}
base_dates <- tibble(
    dates=c(
            seq(as.Date("2000-08-20"), as.Date("2000-12-10"), by="day"),
            seq(as.Date("2001-08-20"), as.Date("2001-12-10"), by="day"),
            seq(as.Date("2002-08-20"), as.Date("2002-12-10"), by="day"),
            seq(as.Date("2003-08-20"), as.Date("2003-12-10"), by="day"),
            seq(as.Date("2004-08-20"), as.Date("2004-12-10"), by="day"),
            seq(as.Date("2005-08-20"), as.Date("2005-12-10"), by="day")
        )
    )

base_dates
```

# Create a table of unique combinations of all base dates with all 26 oris. This ensures that
# dates where no offenses were recorded are included in future stats. 
```{r}
unique_combos <- crossing(date=base_dates$dates, ori=ori_mapping$ori)
```

# Total offense counts (assault/vandalism) by day for each of the 26 oris
```{r}  
final_offense_count <- left_join(unique_combos,offense_count,by=c("ori","date")) %>%
    replace_na(list(assault=0,vandalism=0))

final_offense_count
```

# REPRODUCING TABLE 1
```{r}
appendix_table <- tibble(
    stat=c("mean","sd","0.25q","0.50q","0.75q","0.90q"),
    assault=c(mean(final_offense_count$assault),
        sd(final_offense_count$assault),
        quantile(final_offense_count$assault, probs=0.25),
        quantile(final_offense_count$assault, probs=0.5),
        quantile(final_offense_count$assault, probs=0.75),
        quantile(final_offense_count$assault, probs=0.9)),
    vandalism=c(mean(final_offense_count$vandalism),
        sd(final_offense_count$vandalism),
        quantile(final_offense_count$vandalism, probs=0.25),
        quantile(final_offense_count$vandalism, probs=0.5),
        quantile(final_offense_count$vandalism, probs=0.75),
        quantile(final_offense_count$vandalism, probs=0.9))
)

appendix_table
```

# REPRODUCING TABLE 2
## Recreate table 2 that creates a weekday summary of games and observations

```{r}
# Based on the initial schedule file,
# we filter out the home or away team that played daily

game_week <- filtered_sched %>%
    mutate(day_of_the_week=wday(game_date, label = TRUE, abbr = FALSE))%>% 
    group_by(day_of_the_week) %>% 
    summarize(num_games=n())
    
game_week2 <- game_week %>% 
  bind_rows(tibble(day_of_the_week = "Total", num_games = sum(game_week$num_games)))
```

```{r}
# Observations represent the number of times each days were observed 

observations <- final_offense_count %>%
    mutate(day_of_the_week=wday(date, label = TRUE, abbr = FALSE)) %>%
    group_by(day_of_the_week) %>%
    summarize(observation_count=n())
  
observations2 <- observations %>%
    bind_rows(tibble(day_of_the_week = "Total", observation_count = sum(observations$observation_count)))
```

```{r}
table_2 <- left_join(game_week2,observations2)
table_2
```